---
title: "Global_Terrorism_Data_Analysis_And_Visualization"
output: html_document
date: "2025-05-09"
---

#ROLE 1: Navya - Data wrangling 
#ROLE 2: Pooja - Data Analysis 
#ROLE 3: Rajesh - Data Visualization

#Role 1: Filter and group data by target type.

```{r}
# Data Wrangling
library(tidyverse)
# Import the dataset 
terrorism_data<-read.csv("globalterrorismdb_0718dist.csv")
```

```{r}
# Initial Data Exploration
# View the first few rows to get an idea of the structure
head(terrorism_data)

# Display the structure of the dataset: types of each column
str(terrorism_data)

# Summary statistics to understand distributions and detect possible data quality issues
summary(terrorism_data)

```
```{r}
# Check for missing values in the target type column (as part of validation)
sum(is.na(terrorism_data$targtype1_txt))
```
### Observations: As per the task for group 4, it is observed that there are no missing values for the target type1 column.
```{r}
# Since no missing values were observed in 'targtype1_txt', we can proceed
# Select relevant columns and order by target type code
terrorism_data_ordered <- terrorism_data %>% 
  select(targtype1, targtype1_txt) %>%  # Keep only target type columns
  arrange(targtype1)  # Arrange the data by the coded target type

```

```{r}
# View the ordered dataset
print.data.frame(head(terrorism_data_ordered, 10))  # Display first 10 rows of the cleaned and arranged dataset

# View the number of rows and columns in the selected data
nrow(terrorism_data_ordered)  # Total rows
ncol(terrorism_data_ordered)  # Total columns

```
#Role 2 Analysis Questions: Count the number of attacks on each target type

```{r}
# Load dplyr
library(dplyr) # Loads the dplyr package, which is used for data manipulation using piping and tidy verbs like group_by and summarise.

# Group by count using dplyr
number_of_attack_per_target_type <- terrorism_data_ordered %>% # Takes the dataset `terrorism_data_ordered`
  group_by(targtype1_txt) %>% # Groups the data by the 'targtype1_txt' column, which represents the type of target (e.g., government, private citizens).
  summarise(total_count = n()) # Summarises the grouped data by counting the number of rows (i.e., attacks) in each group using `n()`.

# Print the result is stored in `number_of_attack_per_target_type`, a data frame showing the number of attacks per target type.
print.data.frame(number_of_attack_per_target_type)
```
#Role 3 - Data Visualization - Bar plot of attacks per target type
```{r}
# Load ggplot2 package for data visualization
library(ggplot2)

# Create a bar plot using the dataset `number_of_attack_per_target_type`
ggplot(number_of_attack_per_target_type, 
       aes(x = reorder(targtype1_txt, total_count),  # Reorder target types by total count
           y = total_count)) +                        # Set y-axis as the total number of attacks
  geom_bar(stat = "identity", fill = "skyblue") +     # Draw bars with actual values (identity), colored skyblue
  theme_minimal() +                                   # Use a minimal theme for a clean look
  labs(                                               
    title = "Number of Attacks on Each Target Type",  # Set plot title
    x = "Target Type",                                # Set x-axis label
    y = "Number of Attacks"                           # Set y-axis label
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels 45 degrees for readability

```

